---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
---


## Loading and preprocessing the data

1. Load the data.

```{r}
if (!file.exists("./activity.csv")) {
    unzip("./activity.zip")
}
activityDS <- read.csv("./activity.csv", header=TRUE, stringsAsFactors=FALSE)
```

2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r}
activityDS$date <- as.Date(activityDS$date)
```

## What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day

```{r}
totalSteps <- aggregate(steps ~ date, data=activityDS, FUN=sum, na.rm=TRUE)
barplot(totalSteps$steps, names.arg=totalSteps$date, xlab="date", ylab="steps")
```

2. Calculate and report the **mean** and **median** total number of steps taken per day

```{r}
meanTotalSteps <- mean (totalSteps$steps)
medianTotalSteps <- median (totalSteps$steps)
```

The mean is **`r format(meanTotalSteps, digits = 7)`** and
the median is **`r format(medianTotalSteps, digits = 7)`**

## What is the average daily activity pattern?

1. Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, 
   averaged across all days (y-axis)
   
```{r}
stepsInterval <- aggregate(steps ~ interval, data=activityDS, FUN=mean)
plot(stepsInterval, type="l")
```

2. Which 5-minute interval, on average across all the days in the dataset,contains the maximum number of steps?

```{r}
stepsInterval$interval[which.max(stepsInterval$steps)]
```

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with `NA`s)

```{r}
sum(is.na(activityDS))
```

2. Devise a strategy for filling in all of the missing values in the dataset. 
   The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, 
   or the mean for that 5-minute interval, etc.
  
  To filling missing values, I choose to replace the mean value at the same interval across days.
  
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}

fillingNA <- function (data, defaultData) {
  naPosition <- which(is.na(data$steps))
  naReplacement <- unlist(lapply(naPosition, FUN=function(idx) {
    interval <- data[idx, ]$interval
    defaultData[defaultData$interval == interval, ]$steps
  }))
  fillingSteps <- data$steps
  fillingSteps[naPosition] <- naReplacement
  fillingSteps
}

filledActivityDS <- data.frame(
  steps    = fillingNA (activityDS, stepsInterval),
  date     = activityDS$date,
  interval = activityDS$interval)

head(filledActivityDS)
```


4. Make a histogram of the total number of steps taken each day and Calculate and report the **mean** and **median** total number of steps 
   taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing 
   missing data on the estimates of the total daily number of steps?

```{r}
filledTotalSteps <- aggregate(steps ~ date, data=filledActivityDS, FUN=sum, na.rm=TRUE)
barplot(filledTotalSteps$steps, names.arg=filledTotalSteps$date, xlab="date", ylab="steps")

meanFilledTotalSteps <- mean (filledTotalSteps$steps)
medianFilledTotalSteps <- median (filledTotalSteps$steps)
```

* Mean after populate missing values is **`r format(meanFilledTotalSteps, digits = 7)`**  
* Median populate missing values is **`r format(medianFilledTotalSteps,  digits = 7)`**

* Mean before populate missing values is **`r format(meanTotalSteps, digits = 7)`**  
* Median before missing values is **`r format(medianTotalSteps,  digits = 7)`**

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r, cache=TRUE}
dayType <- function(date) {
    if (as.POSIXlt(as.Date(date))$wday == 0 ||  as.POSIXlt(date)$wday == 6) {
        "weekend"
    } else {
        "weekday"
    }
}
filledActivityDS$daytype <- as.factor(sapply(filledActivityDS$date, dayType))
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was creating using simulated data

```{r}
par(mfrow=c(2,1))
for (type in c("weekend", "weekday")) {
    stepsDaytype <- aggregate(steps ~ interval,
                            data=filledActivityDS,
                            subset=(filledActivityDS$daytype==type),
                            FUN=mean)
    plot(stepsDaytype, type="l", main=type)
}
```

---------------------------